---
title: "Regression Models Project"
author: "Robert Tuck"
date: "June 21, 2015"
output: pdf_document
---

### Background

### Data processing
```{r}
car_data <- data.frame(mtcars)
```

### Data Exploration
```{r, echo=FALSE, results='hide'}
summary(car_data)
```
```{r}
car_data$cyl <- factor(car_data$cyl)
car_data$vs <- factor(car_data$vs)
car_data$gear <- factor(car_data$gear)
car_data$carb <- factor(car_data$carb)
car_data$am <- factor(car_data$am,labels=c('Automatic','Manual'))
```

### Regression Model
Perform a regression on mpg as an outcome of all the predictors and then step through the regressions to look for the best model.
```{r, results='hide'}
initialmodel <- lm(mpg ~ ., data = car_data)
bestmodel <- step(initialmodel, direction = "both")
````
```{r, echo=FALSE, results='hide'}
summary(bestmodel)
```
The adjusted R-squared value of 0.84 which is the maximum obtained considering all combinations of variables. From these results we can conclude that more than 84% of the variability is explained by the above model.

### Inference
Perform a t-test to test for transmission type vs mpg.
```{r}
t.test(mpg ~ am, data = car_data)
```
Based on the t-test results, we reject the null hypothesis that the mpg distributions for manual and automatic transmissions are the same.
#### Conclusion

Based on the analysis done in this project, we can conclude that:

- Cars with Manual transmission get 1.8 more miles per gallon compared to cars with Automatic transmission. (1.8 adjusted for hp, cyl, and wt).

- mpg will decrease by 2.5 for every 1000 lb increase in wt.

- mpg decreases negligibly (only 0.32) with every increase of 10 in hp.

- If number of cylinders, cyl increases from 4 to 6 and 8, mpg will decrease by a factor of 3 and 2.2 respectively (adjusted by hp, wt, and am).

### Appendix
#### Data Exploration
```{r}
summary(car_data)
```

#### Regression Model
Summary of bestmodel data
```{r}
summary(bestmodel)
```
Boxplot showing relationships between transmission types
```{r}
boxplot(mpg ~ am, data = car_data, col = "blue", ylab = "miles per gallon")
```

Residual plots of the the bestmodel
```{r}
par(mfrow=c(2, 2))
plot(bestmodel)
```
